---
title: "Assignment 1"
author: "Rory Flemming"
date: "September 14, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1.1  
Using anorexia data from the MASS package:  
```{r}
require('MASS')
names(anorexia)
levels(anorexia$Treat)
typeof(anorexia$Treat)
typeof(anorexia$Prewt)
typeof(anorexia$Postwt)
```
### (a)  
We see from the description that anorexia is a 72x3 data frame. It contains data relevant to the weight changes of young, female anorexia patients in different treatment conditions. The columns are:  
`Treat`: The treatment condition, three levels. "`Contr`" (Control), "`CBT`" (Cognitive Behavioral treatment), "`FT`" (family treatment). These are likely factors, since they appear as a string, but have integer types.  
`Prewt`: Patient weight (lbs.) at the beginning of study period. This is a postive, real number (as a double).  
`Postwt`: Patient weight (lbs.) at the end of the study period. This is a positive, real number (as a double.  
  
### (b)  
```{r}
mean(anorexia$Prewt)
sd(anorexia$Prewt)
mean(anorexia$Postwt)
sd(anorexia$Postwt)

```
  
### (c)  
```{r}
boxplot(anorexia[,2:3])
```

The above plot shows us that almost all quantile values, except for maybe the second quartile, have increased from pre- to post-study. This supports the intuition from the means, that on average, the patient population gained weight over the study.  
  
### (d)  
```{r}
ft = subset(anorexia, anorexia$Treat == 'FT')
ft = ft[c('Prewt','Postwt')]
```
  
### (e)  
```{r}
Postwt = anorexia$Postwt
mPwt = mean(Postwt)
N = length(Postwt)
numer = sum((Postwt - mPwt)^3) / N
denom = (sum((Postwt - mPwt)^2) / N)^(3/2)
skewPostwt = numer / denom
skewPostwt
corxion = sqrt(N*(N-1))/(N-2)
corrctSkewPostwt = corxion*skewPostwt
corrctSkewPostwt
```
### (f)  
```{r}
require(e1071)
skewness(Postwt,type=1)
skewness(Postwt,type=2)
skewness(Postwt,type=3)
```
`type=1` and `type=2` correspond to the uncorrected and corrected formulas in lecture, respectively.  
  
###(g)  
```{r}
mean(ft$Prewt)
sd(ft$Postwt)
mean(ft$Postwt)
sd(ft$Postwt)
hist(ft$Prewt)
hist(ft$Postwt)
t.test(ft$Postwt,ft$Prewt)
boxplot(ft)
```

## Exercise 1.2  
```{r}
www = 'http://www.uvm.edu/~dhowell/methods8/DataFiles/Add.dat'
ADD = read.table(www, header=1)
```
### (a)  
```{r}
N = nrow(ADD)
N_g60 = nrow(subset(ADD,ADD$ADDSC >=60))
P_g60 = N_g60/N
P_g60
```
The probability that a randomly selected person has an ADDSC score greater than or equal to 60 is `r P_g60`.  
### (b)  
```{r}
N_drop = nrow(subset(ADD,ADD$Dropout == 1))
P_drop = N_drop/N
P_drop
```
The probability that a randomly selected person dropped out of school is `r P_drop`.  
### (c)  
```{r}
dropouts = subset(ADD,ADD$Dropout==1)
drop_g60 = subset(dropouts,dropouts$ADDSC >=60)
sens_SC_drop = nrow(drop_g60)/nrow(dropouts)
sens_SC_drop
```
The proportion of dropouts with an ADDSC of 60 or greater is `r sens_SC_drop`.  
### (d)  
```{r}
nondrop = subset(ADD,ADD$Dropout == 0)
ndrop_l60 = subset(nondrop,nondrop$ADDSC < 60)
spec_SC_drop = nrow(ndrop_l60)/nrow(nondrop)
spec_SC_drop
```
The proportion of non-dropouts with an ADDSC less than 60 is `r spec_SC_drop`.  
### (e)  
```{r}
P_ndrop_g60 = (1-P_drop)*(1-spec_SC_drop)/P_g60
P_ndrop_g60
```
The percentage of misclassified students (nondropouts with ADD scores >=60) is `r P_ndrop_g60`.  
### (f)  
```{r}
n_ndrop_g60x = nrow(subset(nondrop,nondrop$ADDSC >= 60))
p_ndrop_g60 = (n_ndrop_g60x / nrow(ADD))/P_g60
p_ndrop_g60
```
### (g)  
Misdiagnoneses occur because the nondropout population is so much larger than the drop out population. The threshold score of 60 is within bounds that enough non-drop out population (~13%) get caught in that range with the benefit of higher sensitivity to dropouts (70%).  
  
## Exercise 1.3  
### (a)  
```{r}
lb_IQ1 = mean(ADD$IQ) - sd(ADD$IQ)
hb_IQ1 = mean(ADD$IQ) + sd(ADD$IQ)
n_low1 = nrow(subset(ADD,ADD$IQ < lb_IQ1))
n_high1 = nrow(subset(ADD,ADD$IQ > hb_IQ1))
P_sd1 = (nrow(ADD) - n_low1 - n_high1)/nrow(ADD)
P_sd1

lb_IQ2 = mean(ADD$IQ) - 2*sd(ADD$IQ)
hb_IQ2 = mean(ADD$IQ) + 2*sd(ADD$IQ)
n_low2 = nrow(subset(ADD,ADD$IQ < lb_IQ2))
n_high2 = nrow(subset(ADD,ADD$IQ > hb_IQ2))

P_sd2 = (nrow(ADD) - n_low2 - n_high2)/nrow(ADD)
P_sd2

```
The percent of IQ scores within 1 and 2 standard deviations from the mean are `r P_sd1` and `r P_sd2`, respectively.  
  
###(b)  
The above percentages are near the same as for a normal distribution (.682, .954).  
  
### (c)  
```{r}
hist(ADD$IQ,15)
```
  
```{r}
set.seed(42)
norm_sample <-rnorm(1000)
hist(norm_sample,15)
```

It appears that the IQ scores appear to be somewhat skewed right, with more values on the mid-lower end than a normal distribution.  
  
### (d)
```{r}
qqnorm(ADD$IQ)
qqline(ADD$IQ)
```
Looking at the QQ plot, you can see that the IQ dataset is a little more right-skewed than a normal distribution. However, it's not very far off.  

## Exercise 1.4  
### (a)  
The `cats` dataset in the `MASS` package is a 144x3 dataset, including sex (`Sex`, M or F), bodyweight (`Bwt`, kilograms) and heart weight (`Hwt`, grams) of a set of adult cats (>2kg).  

### (b)  
Statistical summary of body weight:  
```{r}
mean(cats$Bwt)
sd(cats$Bwt)
skewness(cats$Bwt,type=2)
kurtosis(cats$Bwt,Fasle,type=2)
```
  
Statistical summary of heart weight:  
```{r}
mean(cats$Hwt)
sd(cats$Hwt)
skewness(cats$Hwt,type=2)
kurtosis(cats$Hwt,Fasle,type=2)
```
  
### (c)  
Comparison of Male (1) and female (2) cats' bodyweights:
```{r}
fcats = subset(cats,cats$Sex == 'F'); fcats = fcats[,2:3]
mcats = subset(cats,cats$Sex == 'M'); mcats = mcats[,2:3]
boxplot(mcats[,1],fcats[,1])
```
  
  Comparison of Male (1) and Female (2) cats' heart weights:  
```{r}
  boxplot(mcats[,2], fcats[,2])
```

### (d)  
P(heart weight >= 10 | Female) = ...
```{r}
cats_g10 = subset(cats,cats$Hwt >= 10)
fcats_g10 = subset(cats_g10,cats_g10$Sex == 'F')
p_f_and_g10 = nrow(fcats_g10)/nrow(cats)
p_f = nrow(fcats)/nrow(cats)
p_g10_f = p_f_and_g10 / p_f
p_g10_f
```
  
The probability of a cat's heart weight being greater than or equal to ten grams given that it is a female is `r p_g10_f`.  
  
### (e)  
P(Female | heart weight >= 10)
```{r}
p_g10 =  nrow(subset(cats,cats$Hwt >=10))/nrow(cats)
p_f_given_g10 = p_f_and_g10 / p_g10
p_f_given_g10
```

The probability of  a cat being female, given its heart weight is greater than or equal to 10 is `r p_f_given_g10`.