---
title: "PSY 8814 - Assignment 2"
author: "Rory Flemming"
date: "September 28, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
require(ggplot2)
require(MASS)
require(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 2.1  
This question uses the `diamonds` dataset in `ggplot2`. So we will start with importing `ggplot` and gettiing quick look at the data:
```{r diamonds_summary}
# Look at data summary
data = diamonds
summary(diamonds)
```

### (a,b)
```{r diamonds_scatter}
#Plot relationship of carat-price with color to indicate the color quality of the diamond
ppc <- ggplot(diamonds, aes(x =carat, y =price, color = color)) +
  geom_point() +
  ggtitle("Round Cut Diamonds: Price by Carat") +
  xlab("Carat") +
  ylab("Price (USD)") +
  coord_cartesian(
    xlim = c(0, 5.2),
    ylim = c(min(diamonds$price), max(diamonds$price))
  )
ppc

```

### (c)  
```{r diamonds_clarity}
# Split Carat-Price by clarity
ppc + facet_wrap(vars(diamonds$clarity),nrow=2)
```

### (d)  
Generally, higher carat diamonds are higher price. Diamonds with better color and clarity have steeper price curves than ones with worse color or worse clarity.  
  
## Exercise 2.2  
```{r cabbages_summary}
# Look at data summary
data2 = cabbages
summary(data2)
```

### (a)  
``` {r cabbage_head_scatter}
# Plot relationship between HeadWt, VitC
chw <- ggplot(data2, aes(x = HeadWt, y = VitC)) +
  geom_point() +
  ggtitle("Cabbage Head wt vs Vit C") +
  xlab("Head Weight (kg)") +
  ylab("Vitamin C (units)")
chw +
  geom_hline(yintercept=mean(data2$VitC)) + 
  geom_vline(xintercept=mean(data2$HeadWt)) # Add lines for means
```

### (b)  
```{r cult_cabbage_plots}
# Split HeadWt-VitC plots for cultivar level
chw+facet_wrap(vars(data2$Cult),nrow=1)
```
  
### (c)  
```{r cult_headwt_rel, echo=TRUE}
#Subset by cultivar and take the means
c52 = subset(data2,as.character(data2$Cult) == 'c52');
mc52 = mean(c52$HeadWt)
c39 = subset(data2,as.character(data2$Cult) == 'c39');
mc39 = mean(c39$HeadWt)
p_c39gmc52 = nrow(filter(c39,c39$HeadWt>mc52))/nrow(c39)
# Proportion of c39 w/ hw > mean(hw(c52))
p_c39gmc52
#Proportion of all w/ hw b/w the means
p_between = nrow(filter(data2,(data2$HeadWt > mc52) & (data2$HeadWt < mc39)))/nrow(data2)
p_between

```
The proportion of `HeadWt` values from `c39` cabbages that fall above the mean `HeadWt` values from `c52` cabbages is `r p_c39gmc52`. The proportion of all `HeadWt` values that fall between to the means of  `HeadWt` values of the two cultivar levels is `r p_between`.
  
### (d)  
```{r cabbage_date_mod}
# Plot HeadWt-VitC with color to indicate planting date
ggplot(data2, aes(x=HeadWt,y=VitC,color=Date)) +
  geom_point() +
  ggtitle("Cabbage Head wt vs Vit C") +
  xlab("Head Weight (kg)") +
  ylab("Vitamin C (units)") +
  theme_minimal()
```