---
title: "PSY 8814 - Assignment 4"
author: "Rory Flemming"
date: "October 11, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Exercise 4.1  
Jim was interested in anxiety in a particular population with a standard deviation of $\sigma$ = 6 on a well-known anxiety scale. The N = 9 individuals had the following scores: 36, 27, 30, 33, 22, 30, 30, 24, 38. Construct a 95% confidence interval using R.  
```{r E4.1}
# Init values
sd = 6 # population sd
N = 9 # sample size
scores = c(36, 27, 30, 33, 22, 30, 30, 24, 38) # our sample data
alpha = .05 # value under tails outside of confidence interval

# calculate sample statistics
xhat = mean(scores) # sample mean
sdhat = sd/sqrt(N) # standard error

# Calculate the interval boundaries
low_bound = xhat + qnorm(alpha/2) * sdhat
high_bound = xhat + qnorm(1 - alpha/2) * sdhat

CI = c(low_bound,high_bound)
CI
```
The 95% confidence interval is between ``[`r CI[1]`,`r CI[2]`]``.  

## Exercise 4.2  
Based on the same population standard deviation and sample as in Exercise 4.1 , please use R to construct a 80% confidence interval.  
```{r E4.2}
alpha = .2 # new alpha
# Calculate the interval boundaries
low_bound = xhat + qnorm(alpha/2) * sdhat
high_bound = xhat + qnorm(1 - alpha/2) * sdhat

CI = c(low_bound,high_bound)
CI

```
The 80% confidence interval is between ``[`r CI[1]`,`r CI[2]`]``.  

## Exercise 4.3  
Please compare between the results you get from Exercise 4.1 and Exercise 4.2, then draw a conclusion on the relationship between confidence level and the width of confidence interval.  

**Answer:** Higher levels of confidence result in tighter bounds on the interval.  

## Exercise 4.4  
Based on the same population standard deviation and sample as in Exercise 4.1, is it *likely* that this sample is draw from a population with mean $\mu$ = 22? Please use $\alpha$ = 0.05. Please explain your answer.
```{r E4.4}
#Init variables
mu = 22
alpha = .05

# Construct the 1-alpha CI for the population distribution with mu =22
H0_low_bound = mu + pnorm(alpha/2,mu,sd)*sd
H0_high_bound = mu + pnorm(1-alpha/2,mu,sd)*sd

if ((xhat<H0_low_bound) || (xhat>H0_high_bound)){
  likely = FALSE
}  else{
  likely = TRUE
}
likely
```
Constructing the null hypothesis distribution and using $\alpha$ = 0.05 as the criterion, I would asses the statement: "it is likely that this sample is drawn from a popilation with mean, $\mu$ = 22" as `r likely`.