---
title: "Lecture 18 Activity"
author: "Rory Flemming"
date: "November 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(tidy_ops=list(width.cutoff=60),tidy=TRUE)
library(dplyr)
library(ggplot2)
```

Import data:  
```{r import}
data=read.csv('sleepdata.csv');
summary(data)
```
Next, calculate some simple statistics:
```{r smpl_stats}
mu = mean(data$rt);
alphas = c(); #init
i=0; #counter
for (k in levels(data$group)){
  i = i+1;
  alphas = c(alphas, mu - mean(data$rt[data$group==k]))
  sigma2_k = length(data$group[data$group==k])*(alphas[i]^2);
}
mu
alphas
residuals = data$rt - mu + alphas[data$group]
residuals
```
Now let's look at the residuals:
```{r resids}
ggplot(NULL,aes(x=1:length(data$rt),y = residuals)) + geom_point() + theme_minimal()
summary(residuals)
```
Residuals look alright. LEt's get the MSB (mean-squared error between groups):  
```{r MSB}
# MSB = var(resid)^2 + sum_k(n_k*alpha_k^2)/K-1
K=length(levels(data$group));
MSB = var(residuals) + sum(sigma2_k)/(K-1)
MSB
```
Let's look at the ANOVA of the data:
```{r anova}
aov.data = aov(rt~group,data)
summary(aov.data)
```
And now a linear (structural) model:
```{r lm}
lm.data = lm(rt~group, data)
summary(lm.data)
```
## Questions  
1. What estimates/statistics are the same in both outputs?  
*Both output F-statistics*
2. What is the intercept term equal to in the second output?  
The intercept give the smallest 